\textsf{Are large language models superhuman chemists?}

### 2024-10-29 src/tex/appendix.tex
chore: change the example for reasoning

While adding questions from existing benchmarks might seem to be another good source of semi-automatically generated data, we prioritized the diversity of the data and avoided data contamination while addressing the guidelines above and in \Cref{sec:desired-properties}.
However, even though we decided not to include questions from other previously published chemistry-focused benchmarks into the \chembench corpus, the framework is flexible enough to be readily extended with questions from other benchmarks.

@@

    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\

@@

            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\

@@

            What is the oxidation number of the metal in the compound \ce{[ZrF7]^{3-}}\\

@@

            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose \ce{HF} over \ce{HCl} for this purpose? \\

@@

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different \enquote{requires} skills, i.e. Reasoning and Calculation.} \\
    \label{tab:chembench_corpus_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following salts is an acidic salt? \\
            A. \ce{NH4Cl} \\
            B. \ce{Na2CO3} \\
            C. \ce{NaH2PO4} \\
            D. \ce{Zn(OH)Cl} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
            A. By atomic force microscopy \\
            B. By transform infrared spectroscopy \\
            C. By measuring a photoluminescence spectrum \\
            D. By absorption spectroscopy \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}

In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers. The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement

@@

    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
    \label{tab:chembench_corpus_models_failed_topic}

@@

            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\

@@

            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
            A. DESI-mass spectrometry \\
            B. XPS \\
            C. atomic force microscopy \\
            D. scanning tunneling microscopy \\
            E. secondary ion mass spectrometry \\

@@

            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\

@@

            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\

@@

            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\

@@

            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
            A. \ce{SCl6} does not exist because of steric effects \\
            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
            A. All dimension of the material must be smaller than exciton Bohr radius \\
            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
            C. The movement od charge carriers is limited in all spatial dimensions \\
            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
            What makes acrylates more reactive than typical hydrocarbons? \\
            A. The carbonyl group \\
            B. The double bond \\
            C. The hydroxyl group \\
            D. The vinyl group \\
            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
            Why are polar molecules more soluble in water than non-polar molecules? \\
            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
            B. The entropy of the system increases when non-polar molecules dissolve in water \\
            C. Water has a dipole and can easier isolate polar molecules from each other \\
            D. Water makes stable dipole-dipole interactions with polar molecules \\
            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
            A. Broader product diameter distribution than pellets or extrudates \\
            B. Facilitates reactor upscaling \\
            C. It is also called snow-ball \\
            D. Not suitable to prepare egg-shell catalysts \\
            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
            What determines the flammability of polyols? \\
            A. Their chemical structure \\
            B. Their low volatility \\
            C. Their odor threshold \\
            D. Their reaction with acids \\

@@

To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them. The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
    \label{tab:chembench_corpus_models_failed_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following are true for phosphoric acid? \\
            A. Is a weak acid, so it is not toxic \\
            B. Should never be tasted \\
            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            Which of the following elements has the largest atomic radius? \\
            A. \ce{Fr} \\
            B. \ce{He} \\
            C. \ce{Li} \\
            D. \ce{Rb} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}. What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}

>>>>

@@

    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}

@@

            A. [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
            B. [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\

@@

            What is the oxidation number of the metal in the compound \ce{[ZrF_{7}]3-}? \\

@@

            For NMR analysis, you need to digest the MOF in a strong acid to remove the linker and leave the metal clusters intact. Why would one choose HF over \ce{HCl} for this purpose? \\

@@

@@

    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different \enquote{requires} skills, i.e. Reasoning and Calculation.}
    \label{tab:chembench_corpus_cognitive}

@@

            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\

@@

@@

            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            You make the following experimental observations: The sample consisted of small colorless/white crystals, wine-red transparent crystals, turquoise transparent crystals, and violet flakes. Most components dissolved in cold water, with turquoise crystals dissolving upon heating. Dimethylglyoxime produced a raspberry-red color with dissolved turquoise crystals. Violet flakes turned blue on contact and showed red/violet colors in flame tests. An oxidation melt turned green. No red coloration occurred with \ce{KSCN}, but a blue organic phase formed with acetone. Potassium ferrocyanide showed no changes. \ce{KOH} produced a brown/black precipitate. Alizarin-S testing resulted in red coloration after acidification. Heating with cobalt nitrate produced a bluish color without luster. Lead dioxide and sulfuric acid oxidation formed a violet solution. No changes occurred in chromate oxidation or titanium detection attempts. \ce{BaCl2} formed precipitates in both original solution and soda extract. Lunge's reagent showed red coloration around zinc granules. Silver nitrate formed white precipitates that dissolved in \ce{NH3} and reappeared with \ce{HNO3} acidification. The separation procedure began with the aqueous solution being treated with \ce{(NH4)2CO3} solution until a persistent turbidity appeared, followed by addition of dilute \ce{HCl} until the turbidity disappeared. The pH was then adjusted to 5-6 with \ce{NH3} (2 M) and heated on a water bath. Three spatula tips of urotropin were added, resulting in a white, rather flocculent precipitate. This precipitate was centrifuged off and washed with warm water. In the Urotropin group, the residue was dissolved in dilute \ce{HCl} and diluted with water. An alkaline precipitation was performed, resulting in a brown coloration even after repeated washing of the urotropin precipitation. This precipitate dissolved in 2 M \ce{HCl}, but showed no change when treated with either \ce{KSCN} or potassium ferrocyanide. In the \ce{(NH4)2S} group, addition of dimethylglyoxim produced a very flocculent red precipitate that rose to the top of the test tube. When \ce{H2S} was introduced to the centrifugate, a black precipitation occurred. The precipitates did not completely dissolve in \ce{HCl}. Treatment with \ce{KSCN} resulted in a blue organic phase when layered with acetone. An alkaline precipitation produced a brown/black residue. No precipitation was observed when \ce{H2S} was introduced to the supernatant liquid after the alkaline precipitation.  What ions are in the sample? \\
            A. \ce{Al^{3+}} \\
            B. \ce{Co^{2+}} \\
            C. \ce{Cr^{3+}} \\
            D. \ce{Fe^{3+}} \\
            E. \ce{Mn^{2+}} \\
            F. \ce{Ni^{2+}} \\
            G. \ce{Zn^{2+}} \\
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\

@@

            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\

@@

            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\

@@

            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\

@@

In analyzing the performance of state-of-the-art models on the \chembench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers. The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of some questions failed for all the state-of-the-art models.} The table shows examples of the \chembench corpus for which all models considered in \Cref{fig:human_vs_models_bar} fail to answer correctly.}
    \label{tab:chembench_corpus_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            Which of the following compounds are capable of producing hazardous explosive peroxides? \\
            A. Cyclohexene \\
            B. Cyclooctene \\
            C. Decalin \\
            D. Ethyl ether \\
            E. Isopropyl ether \\
            F. Para-dioxane \\
            G. Tetralin \\
            \midrule
            What is the oxidation number of hydrogen in the reactants of Sabatier reaction? \\
            \midrule
            How does a non-competitive inhibitor influence the enzyme kinetics of a reaction in the Lineweaver-Burk-plot when compared to a competitive inhibitor? \\
            A. The intersection with y-axis shift to higher numbers \\
            B. The intersection with y-axis shift to lower numbers \\
            C. The slope decreases \\
            D. The slope increases \\
            \midrule
            Poly(lactic acid) (PLA) can be broken down enzymatically to lactic acid monomer and compound B (a dimer of lactic acid). A plastic sample degraded to 1.044 g, and titration showed it needed \SI{19.40}{cm^{3}} of \SI{0.100}{mol dm^{-3}} \ce{NaOH} to neutralize. What is the number of moles of acid in the sample? \\
            \midrule
            What is hydrofluoric acid used for? \\
            A. For cleaning facades \\
            B. For etching glass \\
            C. For filling accumulators \\
            D. For soldering \\
            \midrule
            How many chiral centers are present in the molecular structure of citric acid? \\
            \midrule
            What is the sum of the stoichiometric coefficients on the left side of the redox reaction equation: \ce{Cd(NCS)2 + KNO3 -> K2CO3 + CO + SO2 + N2 + CdO}? \\
            A. 10 \\
            B. 14 \\
            C. 18 \\
            D. 23 \\
            E. 9 \\
            \midrule
            The ligand-field splitting energy (LFSE) is computed as $LFSE=-0.4\cdot Dq \cdot n+0.6 \cdot Dq \cdot m$ with $n$ being the number of electrons in t2g d orbitals ($\mathrm{d}_{xz}$, $\mathrm{d}_{yz}$ and $\mathrm{d}_{xy}$), $m$ the number of electrons in eg d orbitals ($\mathrm{d}_{x^2-y^2}$ and $\mathrm{d_{z^2}}$), and Dq the ligand field splitting parameter. What is the value of LFSE in terms of Dq for the low-spin case in the metal complex tetramethylamine diaquacopper(II) ion? \\
            \midrule
            In phenol's mass spectrum, how does the intensity of the m/z 95 peak compare to the molecular ion peak at m/z 94? \\
            A. 10.0\% \\ 
            B. 6.6\% \\
            C. 7.0\% \\
            D. 94\% \\
            \bottomrule
        \end{tabularx}
    }
\end{table}
